Gap formation and stability in non-isothermal self-gravitating discs

### 2014-05-30 paper/model.tex
tex updates

\subsection{Simulation procedure}
The disc is evolved without a planet for $t\in[0,20]P_0$. The planet potential is smoothly switched on
between $t\in[20,30]P_0$. In a separate set of calculations the planet
potential is turned off after $t>t_\mathrm{off}$, immediately  
afterwhich the surface density, energy density and velocity fields are
azimuthally averaged and then subject to random surface density  
perturbations of magnitude $10^{-2}$.

\subsection{Diagonistic measures}
%% The Toomre parameter is 
%% \begin{align}\label{toomre}
%% Q \equiv \frac{\kappa c_s}{\pi G \Sigma},
%% \end{align}
%% where $\kappa^2 = r^{-3}\p(rv_\phi)/\p r$ is the square of the epicyclic frequency and $c_s = \sqrt{\gamma} \ciso $ is the adiabatic sound-speed. 
%% Note that Eq. \ref{toomre} differs from the Keplerian Toomre parameter used to label the disc models. 
%The potential vorticity (PV), or vortensity, is 
%\begin{align}
%\eta \equiv \frac{\kappa^2}{2\Omega\Sigma}.
%\end{align}
%Note that $\eta$ is related to $Q$ through $\kappa$.

According to previous 2D studies, the key quantitiy in describing the
stability of radially structured non-barotropic discs is the
\emph{generalized vortensity} (GV): 
\begin{align}
  \tilde{\eta}\equiv \left(\frac{\kappa^2}{2\Omega\Sigma}\right)S^{-2/\gamma}, 
\end{align}
where $S \equiv p/\Sigma^\gamma$ is defined to be the entropy. The
quantity in brackets, $\eta\equiv \kappa^2/2\Omega\Sigma$, is the
usual vortensity or potential vorticity (PV).

%It will be convenient to analyse results in terms of perturbations with respect to a smooth background. Let 
%\begin{align}
%\Delta f = \frac{f-\avg{f}}{\avg{f}}
%\end{align}
%denote the perturbation of a flow quantity $f$

>>>>

\subsection{Simulation procedure}
The disc is evolved without a planet for $t\in[0,20]P_0$. The planet potential is smoothly switched on
between $t\in[20,30]P_0$.
%%  In a separate set of calculations the planet
%% potential is turned off after $t>t_\mathrm{off}$, immediately  
%% afterwhich the surface density, energy density and velocity fields are
%% azimuthally averaged and then subject to random surface density  
%% perturbations of magnitude $10^{-2}$.

\subsection{Diagonistic measures}
%% According to previous 2D studies, the key quantitiy in describing the
%% stability of radially structured non-barotropic discs is the
%% \emph{generalized vortensity} (GV): 
%% \begin{align}
%%   \tilde{\eta}\equiv \left(\frac{\kappa^2}{2\Omega\Sigma}\right)S^{-2/\gamma}, 
%% \end{align}
%% where $S \equiv p/\Sigma^\gamma$ is defined to be the entropy. The
%% quantity in brackets, $\eta\equiv \kappa^2/2\Omega\Sigma$, is the
%% usual vortensity or potential vorticity (PV).

### 2014-05-30 paper/results.tex
tex updates

\section{Gaps in low mass viscous discs}

We first describe results for weakly self-gravitating
discs $Q_0=8$ (giving $M_d=0.015M_*$) so that the gap remains 
gravitationally stable \citep{lin11b}. We also impose a dimensionless
kinematic viscosity $\hat{\nu}=2\times10^{-5}$ to suppress the Rossby
vortex instablity \citep{valborro07}. These parameter values
are typical for disc-planet simulations and yield stable/steady 
gap profiles. This will be useful reference cases to  
understand the unstable cases considered later.

Fig. \ref{lvisc_steady_gap} shows the steady state gap profiles in terms of the relative
surface density and aspect-ratio at $t=200P_0$. Three levels of cooling are considered:
$\tbeta=0.1,\,1.0$ and $10.0$ (i.e. $t_c\Omega_k\simeq2.4,\,24$ and $240$). We refer to them as
fast, moderate and slow cooling cases, respectively.

@@

\begin{figure}
  \includegraphics[scale=.42,clip=true,trim=0cm 1.8cm 0cm 0cm]{figures/compare_profiles_dens020.ps}\\
  \includegraphics[scale=.42]{figures/compare_profiles_h020.ps}
  \caption{Steady-state gap profiles in a low mass viscous disc. The
    surface density perturbation (top) and disc aspect-ratio (bottom)
    are shown as a function of the cooling parameter:  
    $\tbeta=0.1$ (solid, fast cooling), $\tbeta=1$ (dotted,
    moderate cooling) and $\tbeta=10$ (dashed, slow
    cooling). \label{lvisc_steady_gap}}  
\end{figure}

\begin{figure}
  \includegraphics[scale=.42]{figures/compare_profiles_gvort020.ps}
  \caption{Gap structure in a low mass viscous disc, in terms of the
    perturbed generalized vortensity as a function of the cooling
    parameter. \label{lvisc_steady_gvort}} 
\end{figure}

\section{Gaps in massive discs}
We first examine disc models with $Q_o=1.5$. We set the physical
viscosity $\hat{\nu}=10^{-9}$, so that the only energy source is
through shock-heating (via artificial viscosity) and the $\mathcal{C}$
function when $e\Sigma<e_i\Sigma_i$. The numerical resolution is
$N_r\times N_\phi = 512\times 1024$.

%Since we are primarily concerened with gap stability, it is important
%to first examine the gap structured opened by the planet as a function
%of cooling. 
Fig. \ref{gvort1d_q1d5} compares the gap structure in terms of the GV profile for three
levels of cooling: $\tilde{\beta}=0.1$, $\tilde{\beta}=1$ and
$\tbeta=10$. For convenience we will refer to these cases as fast,
intermediate and slow cooling, respectively. The snapshot is taken at
$t=30P_0$, just after the planet is fully introduced.

In terms of the GV profile, the qualitative features of the planetary
gap --- localized GV extrema at the gap edges --- remain unchanged
despite two orders of magnitude difference in the cooling rate.

\begin{figure}
%  \includegraphics[scale=.42,clip=true,trim=0cm 1.8cm 0cm 0cm]{figures/dens1d_q1d5_003_global.ps}\\
%  \includegraphics[scale=.42]{figures/gvort1d_q1d5_003_global.ps}
  \caption{Gap structure in terms of the perturbed surface density
    (top) and perturbed generalized
    vortensity (bottom), as a function of the cooling parameter:  
    $\tbeta=0.1$ (solid, fast cooling), $\tbeta=1$ (dotted,
    intermediate cooling) and $\tbeta=10$ (dashed, slow
    cooling). \label{gvort1d_q1d5}} 
\end{figure}

\begin{figure*}
%  \includegraphics[scale=.55,clip=true,trim=0cm 0cm 0cm
%    0cm]{figures/noniso0_HR_dens004}\includegraphics[scale=.55,clip=true,trim=2.26cm 0cm 0cm
%    0cm]{figures/noniso1_HR_dens004}\includegraphics[scale=.55,clip=true,trim=2.26cm 0cm 0cm
%    0cm]{figures/noniso2_HR_dens004}
  \caption{Gap instability in the heavy disc model, as a function of
    the cooling parameter: $\tbeta=0.1$ (left), $\tbeta=1$ (middle)
    and $\tbeta=10$ (right). The relative surface density
    perturbation is shown. \label{polarxy_q1d5}} 
\end{figure*}

>>>>

\section{Gaps in low mass discs (RL)}

%% We first describe results for weakly self-gravitating
%% discs $Q_0=8$ (giving $M_d=0.015M_*$) so that the gap remains 
%% gravitationally stable \citep{lin11b}. We also impose a dimensionless
%% kinematic viscosity $\hat{\nu}=2\times10^{-5}$ to suppress the Rossby
%% vortex instablity \citep{valborro07}. These parameter values
%% are typical for disc-planet simulations and yield stable/steady 
%% gap profiles. This will be useful reference cases to  
%% understand the unstable cases considered later.

%% Fig. \ref{lvisc_steady_gap} shows the steady state gap profiles in terms of the relative
%% surface density and aspect-ratio at $t=200P_0$. Three levels of cooling are considered:
%% $\tbeta=0.1,\,1.0$ and $10.0$ (i.e. $t_c\Omega_k\simeq2.4,\,24$ and $240$). We refer to them as
%% fast, moderate and slow cooling cases, respectively.

@@

%% \begin{figure}
%%   \includegraphics[scale=.42,clip=true,trim=0cm 1.8cm 0cm 0cm]{figures/compare_profiles_dens020.ps}\\
%%   \includegraphics[scale=.42]{figures/compare_profiles_h020.ps}
%%   \caption{Steady-state gap profiles in a low mass viscous disc. The
%%     surface density perturbation (top) and disc aspect-ratio (bottom)
%%     are shown as a function of the cooling parameter:  
%%     $\tbeta=0.1$ (solid, fast cooling), $\tbeta=1$ (dotted,
%%     moderate cooling) and $\tbeta=10$ (dashed, slow
%%     cooling). \label{lvisc_steady_gap}}  
%% \end{figure}

%% \begin{figure}
%%   \includegraphics[scale=.42]{figures/compare_profiles_gvort020.ps}
%%   \caption{Gap structure in a low mass viscous disc, in terms of the
%%     perturbed generalized vortensity as a function of the cooling
%%     parameter. \label{lvisc_steady_gvort}} 
%% \end{figure}

\section{Gaps in massive discs (MKL)}
%% We first examine disc models with $Q_o=1.5$. We set the physical
%% viscosity $\hat{\nu}=10^{-9}$, so that the only energy source is
%% through shock-heating (via artificial viscosity) and the $\mathcal{C}$
%% function when $e\Sigma<e_i\Sigma_i$. The numerical resolution is
%% $N_r\times N_\phi = 512\times 1024$.

%% %Since we are primarily concerened with gap stability, it is important
%% %to first examine the gap structured opened by the planet as a function
%% %of cooling. 
%% Fig. \ref{gvort1d_q1d5} compares the gap structure in terms of the GV profile for three
%% levels of cooling: $\tilde{\beta}=0.1$, $\tilde{\beta}=1$ and
%% $\tbeta=10$. For convenience we will refer to these cases as fast,
%% intermediate and slow cooling, respectively. The snapshot is taken at
%% $t=30P_0$, just after the planet is fully introduced.

%% In terms of the GV profile, the qualitative features of the planetary
%% gap --- localized GV extrema at the gap edges --- remain unchanged
%% despite two orders of magnitude difference in the cooling rate.

%% \begin{figure}
%% %  \includegraphics[scale=.42,clip=true,trim=0cm 1.8cm 0cm 0cm]{figures/dens1d_q1d5_003_global.ps}\\
%% %  \includegraphics[scale=.42]{figures/gvort1d_q1d5_003_global.ps}
%%   \caption{Gap structure in terms of the perturbed surface density
%%     (top) and perturbed generalized
%%     vortensity (bottom), as a function of the cooling parameter:  
%%     $\tbeta=0.1$ (solid, fast cooling), $\tbeta=1$ (dotted,
%%     intermediate cooling) and $\tbeta=10$ (dashed, slow
%%     cooling). \label{gvort1d_q1d5}} 
%% \end{figure}

%% \begin{figure*}
%% %  \includegraphics[scale=.55,clip=true,trim=0cm 0cm 0cm
%% %    0cm]{figures/noniso0_HR_dens004}\includegraphics[scale=.55,clip=true,trim=2.26cm 0cm 0cm
%% %    0cm]{figures/noniso1_HR_dens004}\includegraphics[scale=.55,clip=true,trim=2.26cm 0cm 0cm
%% %    0cm]{figures/noniso2_HR_dens004}
%%   \caption{Gap instability in the heavy disc model, as a function of
%%     the cooling parameter: $\tbeta=0.1$ (left), $\tbeta=1$ (middle)
%%     and $\tbeta=10$ (right). The relative surface density
%%     perturbation is shown. \label{polarxy_q1d5}} 
%% \end{figure*}
